---
title: "Viz I"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

Import weather data 

```{r}
data("weather_df")
```

First plot

```{r}
#ggplot(data = weather_df, mapping = aes(x = tmin, y = tmax)) + #aes - asthetic, what variables from data you want mapped onto canvas, don't normally see this, see below
  #geom_point()

ggplot(weather_df, aes(x = tmin, y = tmax)) + # more typical code
  geom_point()
```


```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point() #same as code in  chunk above, stylistic choice
```


```{r}
ggp_weather_scatterplot = #when saved in environment, still ggplot object, can add layers
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point() 

ggp_weather_scatterplot 

```


Check that some rows are missing

```{r}
weather_df |> # producing summary information about missingness is very important
  filter(is.na(tmax))

#nanier:: (package - missingess across all columns)
```


## Fancier scatterplots!

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + # color = name, assigns color to each name
  geom_point(alpha = 0.3) + # alpha aes, not in aes b/c it is not mapped, 1 = solid, 0 = see through
  geom_smooth(se = FALSE) # adds smooth line, se = FALSE removes error band around lines, not reliable so remove
```

Where you define aesthetics can matter 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name, size = prcp), alpha = 0.3, size = 0.3) + # if you do this, mapping is not inherited, size = 0.3 size of points, size = prcp variable name, will size points based on variable 
  geom_smooth(se = FALSE)
```

using faceting real quick

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name, size = prcp), alpha = 0.3, size = 0.3) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) # `. ~ name` splits into three columns - cols equal one subgroup/value, `name ~ .` = rows equal one subgroup/value
```

More interesting scatterplot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + # shape = name, maps shape to name, gives different share for each category
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

# Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth())

LA plot:

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr = tmax * (9/5) + 32, 
    tmin_fahr = tmin * (9/5) + 32
  ) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "lm") # method = "lm", fits linear regression

```


## Small things 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + 
  #geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE)

weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + 
  geom_smooth(se = FALSE) +
  geom_point() # added in layers so adding points after smooth makes smooth difficult to see, 
  
```


```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() # separatess x y plane into hexes and shows density of data 
```


```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "purple") # adding color outside of aes to geom point, can also use any color expressable as html hashcode 
```


## Univariate plots 

